---
title: 'Statistical Inference Course Project: Part 2'
author: "Jordan Woloschuk"
date: "8/13/2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview: 


In this project I will analyze the ToothGrowth data contianed withing the
R datasets package. This analysis will include four (4) main steps.

* Load the ToothGrowth data and perform some basic exploratory data analyses
* Provide a basic summary of the data.
* Use confidence intervals and/or hypothesis tests to compare tooth growth by 
supp and dose. 
* State your conclusions and the assumptions needed for your conclusions.


```{r Libraries, include=FALSE}
# Load libraries
library(ggplot2)
```


## Load the ToothGrowth Data

In this section we will load the ToothGrowth data and perform some basic 
exploratory data analyses.

```{r Load}

# Load ToothGrowth
data("ToothGrowth")

# Set the dose as a factor instead of the default num class. This is done for
# future plots.
ToothGrowth$dose <- as.factor(ToothGrowth$dose)


```


## Summary of the ToothGrowth data

In this section we will provide a basic summary of the ToothGrowth data.


```{r Str_data}
# Use str to display the structure of the data
str(ToothGrowth)

```


```{r summary_data}
# Use summary to display a summary of the data
summary(ToothGrowth)

```


```{r head_tail}
# Use head and tail to show the first six (6) and last six(6) rows of data.
head(ToothGrowth)

tail(ToothGrowth)

```