---
title: 'Statistical Inference Course Project: Part 1'
author: "Jordan Woloschuk"
date: "8/13/2019"

output: rmarkdown::github_document
  
---

```{r setup, include=FALSE}
library(knitr)
```

## Overview: 

In this project I will investigate the exponential distribution in R and
compare it with the Central Limit Theorem. 

* The exponential distribution can be simulated in R with rexp(n, lambda) where
lambda is the rate parameter. 

* The mean of exponential distribution is 1/lambda and the standard deviation is
also 1/lambda.

In this project, I will set lambda = 0.2 for all of the simulations.

I will investigate the distribution of averages of 40 exponentials, and will 
conduct a thousand simulations.

```{r Load, include=FALSE}
# Load libraries
library(ggplot2)
```

## Simulations:

We will first set the seed of R‘s random number generator. This is done to 
create simulations or random objects that can be reproduced.

```{r Seed}
# Setting R Seed to 1 (any number would work)

set.seed(1)

```

We will next set the constants that will be used in our simulations.


```{r Constants}

# Number of simulations
numsim <- 1000

# Lamda (lambda is the rate parameter)
lamda <- 0.2

# Number of exponentials
n <- 40

```


We will run the simulation and produce a matix with 40,000 observations 
(n * numsim).

```{r Simulation}
# Run the exponential simulations and populate a numsim * n matrix
Expo_Matrix <- matrix(rexp(n*numsim,lamda),nrow = numsim,ncol = n)

# Calculate the means for each row (create data.frame for ggplot)
Expo_Matrix_Means <- data.frame(expo_means = apply(Expo_Matrix,1,mean))


```

The following plot displays the distributions of exponential means.

```{r Expo Plot}

g_expo <- ggplot(Expo_Matrix_Means, aes(expo_means)) + 
        geom_histogram(color = "navy", fill = "lightblue", binwidth = 0.25) +
        ggtitle("Distribution of Exponential Means") + 
        xlab("Means")+ylab("Frequency")

g_expo

```


## Sample Mean versus Theoretical Mean: 

The theoretical mean (mu) of a exponential distribution of rate lambda is equal
to the 1/lamda.

```{r mu}

# Calculation of mu
mu <- 1/lamda

```

In our example, mu is equal to `r mu`

While, the sample mean (X_bar) is the average of the 1,000 simululations of the 40 sampled exponentials.

```{r X_bar}

# Calculation of X_bar
X_bar <- mean(Expo_Matrix_Means$expo_means)

```

From our simulations, we calculate that the sample mean (X_bar) is equal to `r X_bar`

As expected, the sample mean (X_bar) and the theoretical mean (mu) are very
similar, with an absolute difference of only `r abs(mu-X_bar)`.


## Sample Variance versus Theoretical Variance:

## Distribution: